---
title: "Bertomeu, Ma, and Marinovic (2020, TAR)"
author: "Shunsuke Matsuno"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required Libraries
```{r}
pacman::p_load(tidyverse, ggplot2, patchwork, BertomeuMaMarinovic)
```

# Introduction
- This note simulates BMM (2020).
- Static model is omitted.

# Multi-period Model

## Fixed point of $\Gamma$
- We simulate $\Gamma$ to see numerically that the function has a unique fixed point in the negative region. 
- The function is given by
$$
\Gamma(y) = \frac{1-p_{t}}{p_{t}}\int_{-\infty}^{\infty}\frac{x_{t}}{\sigma_{x}}\phi\left(\frac{x_{t}+y}{\sigma_{x}}\right)\Phi\left(-\alpha x_{t}-\beta\right)dx_{t}.
$$
    - Note that the function involves an integration.
- We use a simple Monte Carlo method to to evaluate $\Gamma(y)$.
    - Fix $y$.
    - By the change of variable $\xi_t = \frac{x_t + y}{\sigma_x}$, the integral is written as
\begin{align*}
\int_{-\infty}^{\infty} & \frac{x_{t}}{\sigma_{x}}\phi\left(\frac{x_{t}+y}{\sigma_{x}}\right)\Phi\left(-\alpha x_{t}-\beta\right)dx_{t}\\
 & =\int_{-\infty}^{\infty}\left(\xi_{t}-\frac{y}{\sigma_{x}}\right)\phi\left(\xi_{t}\right)\Phi\left(-\alpha\sigma_{x}\xi_{t}+\alpha y-\beta\right)\sigma_{x}d\xi_{t}
\end{align*}
    - Write $g(\xi_{t})=\left(\xi_{t}-\frac{y}{\sigma_{x}}\right)\Phi\left(-\alpha\sigma_{x}\xi_{t}+\alpha y-\beta\right)\sigma_{x}$, so the integral is 
$$
\int_{-\infty}^{\infty}g(\xi_{t})\phi(\xi_{t})d\xi_{t}.
$$
    - Therefore, by simulating $\{\xi_t^s \}_{s=1}^{M}$ from $\mathcal{N}(0,1)$, we have
$$
M^{-1}\sum_{s=1}^{M}g(\xi_{t}^{s})\to_{p}\mathbb{E}[g(\xi_{t})]=\int_{-\infty}^{\infty}g(\xi_{t})\phi(\xi_{t})d\xi_{t}.
$$
    - We could use more sophisticated methods of numerical integration, but this simple Monte Carlo method suffices for our purpose.
        - For example, try importance sampling. 
    

- Here is the code for $\Gamma(y)$.
    - Note that $\Gamma$ depends on $(p_{t},\sigma_{x},\alpha,\beta)$, which are treated as parameters.
```{r}
# the definition of Gamma
compute_Gamma
```

- For a set of parameters, we plot $\Gamma(p)$.
    - The parameter values are take from Table 5.
    - By changing parameter values (and the range of $y$), we can verify that a unique fixed point exists in the negative region. 
    
```{r}
# parameters
p <- .3
sigma_x <- .32
alpha <- 2.54
beta <- .67

# simulation
Gamma_simulated <- compute_Gamma(p, sigma_x, alpha, beta)

# plot
ggplot(Gamma_simulated$df, aes(y, Gamma_val)) + 
  geom_point() + 
  stat_function(fun = Gamma_simulated$fun) +
  geom_abline(slope = 1, intercept = 0, colour = "blue") +
  geom_vline(xintercept = 0, lty = 2) + 
  geom_hline(yintercept = 0, lty = 2) + 
  theme_bw()
```


## Simulating Dyunamic Behavior
- In this section, we simulate the dynamic disclosure model.
    - Following the sample of the paper, the panel data is $N \times T$, where 
        - $N$ = 1,000
        - $T$ = 10
- The parameters are taken from the results of the paper.
```{r}
set.seed(2021)

# parameters
p <- .35
sigma_x <- .32
sigma_e <- .74
alpha <- 2.54
beta <- .67
p_11 <- .96    # Phi(1 | 1): prob of theta = 1 to 1
p_01 = .11     # Phi(1 | 0): prob of theta = 0 to 1

# data length
N <- 3 * 1e3
Time <- 10
```



- First, we simulate $\theta$, the value of friction.
    - We draw tge initial value of $\theta_t$ from the Beta distribution. 
    - The evolution is governed by `p_11` and `p_01`.
    - $p_t = \mathbb{E}[\theta_t]$ is also computed.
    
```{r}
# function that generates next theta by FOM
simulate_friction
```


```{r}
# simulation of p and theta
df <- tibble(i = 1:N) %>% 
  mutate(df_p_theta = map(.x = i, .f = simulate_friction,
                           p_11 = p_11, p_01 = p_01, 
                          p_init = rbeta(1, shape1 = 2, shape2 = 5),
                           Time = Time)) %>% 
  unnest(df_p_theta) %>% 
  select(i, t, p, theta)


# overall mean of theta
df %>% pull(theta) %>% mean()
```

- Second, we simulate $x_t$ and observed earnings, $e_t$.
    - Note that $x_t$ may not be observed for certain firms, so we will eliminate the value of $x_t$ for firms with information friction.
    - We simulate $e_t \sim \mathcal{N}(0,\sigma_e^2)$ and then simulate $\tilde{x}\mid e_{t}\sim\mathcal{N}\left(\frac{\sigma_{x}^{2}}{\sigma_{e}^{2}}e_{t},\sigma_{x}^{2}\left(1-\frac{\sigma_{x}^{2}}{\sigma_{e}^{2}}\right)\right)$.
        - See my note for the derivation.

```{r}
df <- df %>% 
  mutate(e = rnorm(N*Time, sd = sigma_e)) %>% 
  mutate(x = rnorm(N*Time, mean = (sigma_x^2)/(sigma_e^2) * e,
                           sd   = sigma_x^2 * (1 - (sigma_x^2)/(sigma_e^2))))
```


- Next, we delete the private signal $x_t$ if the manager is informed ($\theta_t = 0$).
```{r}
df <- df %>% 
  mutate(x = if_else(theta == 0, x, NA_real_))
df
```

- We then compute the utility of disclosing, which is given by
$$
u_t(d_t = 1 \mid x_t) - u_t (d_t \mid x_t) = \alpha\left\{ x_t+P_{t}^{\textit{nd}}\right\} + \beta + \varepsilon_{t}.
$$
    - Effectively, $u_t (d_t \mid x_t)$ is normalized at zero.
- To compute the utility, we have to compute the non-disclosure price, $P_t^{nd}$.
    - To do so, note that we have to derive the fixed points of $\Gamma(y;p_t)$.
    - Following the paper, we estimate $\Gamma(y; p_t)$ by the spline method.

```{r}
compute_fixedpt_Gamma
```

- When $p$ is close to zero, the fixed point algorithm fails because $p/(1-p)$ is too large.

```{r}
# compute fixed points for each p
df_Pnd <- tibble(p = ((1:20)/21)^2) %>% 
  mutate(P_nd = map_dbl(.x = p, .f = compute_fixedpt_Gamma,
                        sigma_x = sigma_x, alpha = alpha, beta = beta,
                        max_iter = 500)) 
Pnd_splined <- splinefun(x = df_Pnd %>% pull(p),
                         y = df_Pnd %>% pull(P_nd))

# plot the estimate P_nd as a function of p
qplot(p, P_nd, data= df_Pnd) +
  stat_function(fun = Pnd_splined) +
  theme_bw()

# Compute P_nd for each p
df <- df %>% 
  mutate(P_nd = map_dbl(.x = p, .f = Pnd_splined))
```

- Now we can compute the utility of disclosure and thus the disclosure decision.
```{r}
df <- df %>% 
  mutate(u = alpha * (x - P_nd) + beta + rnorm(N*Time)) %>% 
  mutate(d = if_else(theta == 1 | u < 0, 0, 1))
df
```

- We delete the value of $x_t$ for non-disclosure firms.
```{r}
df <- df %>% 
  mutate(x = if_else(d == 1, x, NA_real_))
df
```


- We look at the summary statistics and if that matches to Table II of the original paper. 

```{r}
df %>% 
  select(-i, -t) %>% 
  psych::describe(skew = FALSE) %>% 
  select(-vars)
```

- Finally, we drop variables that are not observed.
    - In the next section, we check if the estimation procedure can recover the parameters.
    
```{r}
df_observed <- df %>% 
  select(i, t, e, x, d)
df_observed

# save
save(df_observed, file = here::here("./middle/df_observed.RData"))
```

# Model Identification
- In this section, we replicate Figure 4.
- Simulating observed data for a set of parameters, i.e., the procedure of the previous section, is summarized in  `simulate_model.R`.
- Since each simulation takes a couple of minutes, we will use paralellization. 
    - On my computer (Core i7-11370H 3.30GH, 32GB RAM), one simulation of the model takes about 30 seconds.

- Density of disclosed information for each $\alpha$ (Top left)
```{r}
# parameters
alpha_vec <- c(0, 1, 5, 20)

# register cores
library(doParallel)
cl <- parallel::makePSOCKcluster(4)
doParallel::registerDoParallel(cl)

# simulate data for each alpha
obj = foreach (i = seq_along(alpha_vec),
               .packages = c('dplyr', 
                             'purrr', 
                             'tidyr',
                             'BertomeuMaMarinovic')) %dopar% {
                 BertomeuMaMarinovic::simulate_model(
                   alpha_vec[i], 
                   beta, sigma_x, sigma_e, p_11, p_01, N = 5*1e3, Time = 20
                 )
               }

# kill finished tasks
parallel::stopCluster(cl)

# combine data
temp <- list()
for(k in seq_along(alpha_vec)){
  temp[[k]] <- obj[[k]] %>% 
    mutate(alpha = alpha_vec[k])
}
df_1 <- bind_rows(temp)
rm(obj)

# plot
df_1 %>% 
  filter(d == 1) %>%     # only disclosing sample
  select(x, alpha) %>% 
  ggplot(aes(x, group = factor(alpha))) + 
  geom_density(aes(colour = factor(alpha)),
                   size = 1,
                   alpha = .8) +
  theme_bw()
```

- Probability of disclosure for each $\beta$ and $\alpha$
```{r}
# parameters
param_table <- 
  expand_grid(alpha = seq(0, 5, .5),
              beta = -2:2) %>% 
  as.matrix()

# register cores
cl <- parallel::makePSOCKcluster(8)
doParallel::registerDoParallel(cl)

# simulate data for each alpha
obj <-  foreach (i = 1:NROW(param_table),
               .packages = c('dplyr',
                             'purrr',
                             'tidyr',
                             'BertomeuMaMarinovic')) %dopar% {
                 BertomeuMaMarinovic::simulate_model(
                   param_table[i,1], param_table[i,2],
                   sigma_x, sigma_e, p_11, p_01, N = 1e4, Time = 10
                 )
               }

# kill finished tasks
parallel::stopCluster(cl)

# combine data
temp <- list()
for(k in 1:NROW(param_table)){
  temp[[k]] <- obj[[k]] %>%
    mutate(alpha = param_table[k,1],
           beta = param_table[k,2])
}
df_2 <- bind_rows(temp)

plot
df_2 %>%
  group_by(alpha, beta) %>% 
  summarise(d = mean(d),
            .groups = "drop") %>% 
  ggplot(aes(x = alpha, y = d, group = factor(beta))) +
  geom_point(aes(colour = factor(beta)),
                   size = 1) +
  geom_line(aes(colour = factor(beta))) +
  theme_bw()
```

- 

# Estimation
- Using the simulated observed data, `df_observed`, we estimate the parameters.
- Several details that are not (probably) not mentioned in the article.
    - Estimation of $\Phi(k_0)$ and $\Phi(k_1)$.
        - I guess they used MLE, but how did they decide the initial value of $p_t$?
    - The computation of $L_t^0$.
        - Computing $v_t$ involves numerical integration. How did they do that?

- In this note, the estimation is conducted by the following steps:
1. Estimate $\sigma_e$ from the observed earnings.
2. Estimate the rest of the parameters using MLE.
    - For each set of parameters, we compute $P^{nd}$ and then compute the likelihood function. 
    - The initial value of $p_t$ is drawn from the beta distribution.

- `compute_likelihood` computes the likelihood.
    - To compute likelihood for a set of parameters, it takes about 10 seconds.
```{r}
# data
load(here::here("./middle/df_observed.RData"))
param <- c(alpha = 2.54, beta = .67, sigma_x = .32, p_11 = .96, p_01 = .11)
compute_likelihood(param, df_observed)
```

- Optimization
```{r}
# # make cluster
# library(parallel)
# cl <- makeCluster(4)
# setDefaultCluster(cl = cl)
# clusterExport(cl, "df_observed")
# clusterEvalQ(cl, {library(dplyr); library(magrittr); library(purrr);
#                   library(tidyr); library(BertomeuMaMarinovic)})
# 
# # optimization
# res <- optimParallel::optimParallel(
#   param, compute_likelihood, df = df_observed,
#   method = "L-BFGS-B",
#   lower = c(-10, -10, 0.001, 0.001, 0.001),  # param = (alpha, beta, sigma_x, p_11, p_01)
#   upper = c(10, 10, 5, sigma_e_hat, 0.999),  # sigma_e < sigma_x by construction
#   control = list(fnscale = -1, 
#                  trace = 6,
#                  REPORT = 1),
#   parallel = list(loginfo = TRUE, forward = FALSE))
# 
# # stop cluster
# setDefaultCluster(cl=NULL); stopCluster(cl)
# 
# res
```

- Finally, we check the identification by plotting the likelihood around the true values.

```{r}
plot_list <- list()
for(i in 1:5){
  plot_list[[i]] <- plot_ll(param, df_observed, i)
}
plot_list
```

